---
title: "WUP_04"
author: "Gema Garcia"
date: "Summer 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
thedata<-read.csv(file="GSScsv.csv")
```

```{r eval=FALSE}
thedata <- thedata %>%
mutate(across(where(is.character), as.factor))
```


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->


We looked at the data to try and see what kind of data we had. We copied some codes from writeup 3 and changed the character variables to better fit our data here. It then gave us numerical data and we now have a factor variable called DEGREE.So we deducted we have a numerical ~ categorical problem.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
ggplot(thedata, aes(x = DEGREE, y = WORDSUM)) +
geom_boxplot(fill = "skyblue", color = "darkblue", outlier.shape = NA) +
theme_minimal() +
labs(x = "DEGREE", y = "WORDSUM") +
theme(plot.title = element_text(hjust = 0.5)) +
ggtitle("Fancy Box PLot")
```


We used chatGBT to get our box plot to help us visually see the data. LT high school does not overlap with Junior or Graduate college, therefore showing us it is very different schooling. Same goes with graduate not overlapping with high school, they have different schooling. While some overlap, others do not and that is how we can spot different schooling. 



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
favstats(WORDSUM~DEGREE,data=thedata)
```


The average score on the WORDSUM test for graduates was a little over 10. There's a difference between the average score in Graduate school and the average score in Junior college. Which then shows us there is a difference for schooling between each school or college. 



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
model1 <- aov(WORDSUM~DEGREE,data=thedata)
summary.aov(model1)
```


There is a significant difference between the schooling of all five degrees.



```{r}


# Perform one-way ANOVA
model <- aov(WORDSUM ~ DEGREE, data = thedata)

# Conduct post-hoc test (Tukey"s HSD)
posthoc <- TukeyHSD(model)

# Print the results
print(posthoc)

```


Based on the chart, Lt high school and Graduate are the two degrees that are different from one another.We can also see high school is different than graduate.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


We used chatgpt to help us get the box plots to see how each degree has the same or different schooling. We also used it for the inferential results and saw the averages which helped us determine just how different each degree is. We had some issues with chatgpt not wanting to give us the code we wanted at first but when we kept asking it more questions to give us the code we wanted, it did. 
